---
title: "Analysis of Subcortical Volume Measures Between Random Groups - Comparing Regression Models With and Without Propensity Weights"
author: "Melissa Nunez"
date: "8/24/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r Data Preparation, include = FALSE}
library(gridExtra)
library(tableone)
library(survey)
library(nlme)
library(knitr)
library(kableExtra)
library(dplyr)
library(ggplot2)

study58 <- read.csv("/Users/user/Dropbox/Melissa/Study58/Study58_Merged_NeuroPreAndPost.csv")

study58_rg1 <- study58 %>%
  filter(random_group == 1)
  
study58_rg2 <- study58 %>%
  filter(random_group == 2)

```

##Baseline Differences in Subcortical Volume Measures Between Random Groups. 

For a series of subcortical volume measures, both t-tests and the wilcoxon test were used to test for differences between the randomization groups The following table presents the mean, standard deviation, t-test p-value, and wilcoxon p-value for differences. The final column, "P.Value", gives the p-value for the random_group variable of the following regression model: sub_cor_measure_baseline = random_group + np_gender + IntraCranialVolume.x + np_age where sub_cor_measure_baseline is the subcortical volume measure being analyzed.

```{r Random Group Baseline Comparison, echo = FALSE, warning = FALSE}
# Function calculates mean, standard deviation, checks constant variance, and then runs ttest and wilcoxon test ( 2 variables) .

 run.tests<-function(yvar){

# Mean and Standard Deviation for random groups 1 & 2   
rg1_mean <- mean(study58_rg1%>%select(yvar)%>%unlist(), na.rm = TRUE)
rg2_mean <- mean(study58_rg2%>%select(yvar)%>%unlist(), na.rm = TRUE)
   
rg1_sd <- sd(study58_rg1%>%select(yvar)%>%unlist(), na.rm = TRUE)
rg2_sd <- sd(study58_rg2%>%select(yvar)%>%unlist(), na.rm = TRUE)
  
# Check equality of variance for ttest
eqvar_pval=var.test(study58_rg1%>%select(yvar)%>%unlist(),study58_rg2%>%select(yvar)%>%unlist(), alternative = "two.sided")$p.value

# T-test - p.value
ttest_pval=t.test(study58_rg1%>%select(yvar)%>%unlist(),study58_rg2%>%select(yvar)%>%unlist(), var.equal=(eqvar_pval >.05), paired=FALSE)$p.value

# Wilcoxon test - p.value
Wilcoxtest_pval=wilcox.test(study58_rg1%>%select(yvar)%>%unlist(),study58_rg2%>%select(yvar)%>%unlist(), alternative = "two.sided")$p.value
    
# Extract p.value for variabe of interest while adjusting for variables
regmod <- anova(lm(study58%>%select(yvar)%>%unlist() ~ random_group + np_gender + IntraCranialVolume.x + np_age, data = study58))[1,5]
    varname <- gsub('.{2}$', '', yvar)
    
return(c(varname, round(rg1_mean, 2), round(rg2_mean, 2), round(rg1_sd, 2), round(rg2_sd, 2), round(Wilcoxtest_pval, 4), round(ttest_pval, 4), round(regmod,4) ))
  
 }
 
# Apply function 
rg_summary <- data.frame(do.call(rbind,lapply(names(study58_rg1)[c(50:65,68:79,82:83,90:94)],run.tests)))

colnames(rg_summary) <- c("Variable", "RG1 Mean", "RG2 Mean", "RG1 SD", "RG2 SD", "Wilcoxon Test", "T.Test", "P.Value")

sigvars <- rg_summary %>%
  filter(as.numeric(as.character(P.Value)) <= 0.05) %>%
  select(Variable) %>%
  lapply(paste,".x", sep = "")

rg_summary %>%
 kable(format = "latex") %>%
 kable_styling(latex_options = "scale_down")


# Significant variables: Left.Inf.Lat.Vent, Left.Caudate, Right.Caudate, Right.Accumbens.area.
# Possible Significant variables: Left.Putamen
```

## Distribution of Baseline Variables with Significant Difference Between Random Groups
```{r plots of significant variables, echo = FALSE, warning = FALSE}
# Factor random_group
study58$random_group <- factor(study58$random_group)

# Boxplots of significant variables
# Create function for boxplots
boxplots <- function (yvar) {
  ggplot(study58, aes_string(x="random_group", y = yvar, fill = "random_group")) + 
  geom_boxplot(outlier.shape=NA) + #avoid plotting outliers twice
  geom_jitter(position=position_jitter(width=.1, height=0), shape = 1) + 
  labs(title = paste("Distribution of", yvar), x = "Random Group", y = yvar) + 
  guides(fill=FALSE)
}

# Using lapply and my boxplot function
sigplots <- lapply(unlist(sigvars), boxplots)

# Arrange plots of "Left.Inf.Lat.Vent.x", "Left.Putamen.x", "Right.Caudate.x", "Right.Accumbens.area.x" , and "Left.Caudate.x" .
grid.arrange(sigplots[[1]], sigplots[[3]], ncol = 2)
grid.arrange(sigplots[[6]], sigplots[[8]], ncol = 2)
grid.arrange(sigplots[[4]], ncol = 1)
```


## Regression Models With Change Score of Subcortical Volume Measures as Outcome

The following table presents the mean change score for each subcortical volume measurement, stratified by random group, and the standard deviation. Then, a regreesion model was used to acquire p-values for the random_group variable: 
change_score = random_group + sub_cor_measure_baseline + IntraCranialVolume_post + np_age + np_gender where sub_cor_measure_baseline is the subcortical volume measure being analyzed. Each model was ran twice, one with and another without the propensity weights applied, and the p-values after multiple comparison correction are presented as well. 

```{r Change in Subcortical Volume Regression, echo=FALSE}
runglm <- function(xvar) {

yvar <- paste(gsub('.{2}$', '', xvar),".y", sep = "")
change = study58%>%select(xvar)%>%unlist() - study58%>%select(yvar)%>%unlist()
  
meanchange_rg1 = round(mean(study58_rg1%>%select(xvar)%>%unlist() - study58_rg1%>%select(yvar)%>%unlist(), na.rm = TRUE), 2)

meanchange_rg2 = round(mean(study58_rg2%>%select(xvar)%>%unlist() - study58_rg2%>%select(yvar)%>%unlist(), na.rm = TRUE), 2)
  
rg1_sd <- round(sd(study58_rg1%>%select(xvar)%>%unlist() - study58_rg1%>%select(yvar)%>%unlist(), na.rm = TRUE), 2)
rg2_sd <- round(sd(study58_rg2%>%select(xvar)%>%unlist() - study58_rg2%>%select(yvar)%>%unlist(), na.rm = TRUE), 2)
  
regmod <- anova(lm(change ~ random_group + IntraCranialVolume.y + np_age + np_gender, data = study58))[1,5]

regmod_weights <- anova(lm(change ~ random_group + IntraCranialVolume.y + np_age + np_gender, data = study58, weights = study58$IPW))[1,5]
  
varname <- gsub('.{2}$', '', xvar)
  
return(c(varname, meanchange_rg1, meanchange_rg2, rg1_sd, rg2_sd, round(regmod, 4), round(regmod_weights, 4)))
}

pvals=data.frame(do.call(rbind,lapply(names(study58)[c(50:65)],runglm)))

colnames(pvals) <- c("Variable", "RG1 Mean Change", "RG2 Mean Change", "RG1 SD", "RG2 SD", "Pval.wo.weights", "Pval.w.weights")

pvals["Fdr.wo.weights"] = p.adjust(pvals$Pval.wo.weights,method='BH')
pvals["Fdr.w.weights"] = p.adjust(pvals$Pval.w.weights,method='BH')

kable(pvals, format = "html") %>%
  kable_styling(bootstrap_options = "striped", full_width = F) 


```

For this analysis, the p-values presented are of the interaction term between random_group and age at baseline.
Regression Model: change_score = random_group\*np_age + sub_cor_measure_baseline + IntraCranialVolume_post + np_age + np_gender where sub_cor_measure_baseline is the subcortical volume measure being analyzed.
```{r Change in Subcortical Volume Regression w/ interaction, echo=FALSE}
runglm_inter <- function(xvar) {
  
yvar <- paste(gsub('.{2}$', '', xvar),".y", sep = "")
change = study58%>%select(xvar)%>%unlist() - study58%>%select(yvar)%>%unlist()
  
meanchange_rg1 = round(mean(study58_rg1%>%select(xvar)%>%unlist() - study58_rg1%>%select(yvar)%>%unlist(), na.rm = TRUE), 2)

meanchange_rg2 <- round(mean(study58_rg2%>%select(xvar)%>%unlist() - study58_rg2%>%select(yvar)%>%unlist(), na.rm = TRUE), 2)
  
rg1_sd <- round(sd(study58_rg1%>%select(xvar)%>%unlist() - study58_rg1%>%select(yvar)%>%unlist(), na.rm = TRUE), 2)

rg2_sd <- round(sd(study58_rg2%>%select(xvar)%>%unlist() - study58_rg2%>%select(yvar)%>%unlist(), na.rm = TRUE), 2)
  
regmod <- anova(lm(change ~ random_group*np_age + study58%>%select(xvar)%>%unlist() + IntraCranialVolume.y + np_age + np_gender, data = study58))[1,5]

regmod_weights <- anova(lm(change ~ random_group*np_age + study58%>%select(xvar)%>%unlist() + IntraCranialVolume.y + np_age + np_gender, data = study58, weights = study58$IPW))[1,5]
  
varname <- gsub('.{2}$', '', xvar)

return(c(varname, meanchange_rg1, meanchange_rg2, rg1_sd, rg2_sd, round(regmod, 4), round(regmod_weights, 4)))
}

pvals_inter=data.frame(do.call(rbind,lapply(names(study58)[c(50:65,68:79,82:83,90:94)],runglm_inter)))

colnames(pvals_inter) <- c("Variable", "RG1 Mean Change", "RG2 Mean Change", "RG1 SD", "RG2 SD", "Pval.wo.weights", "Pval.w.weights")

pvals_inter["Fdr.wo.weights"] = p.adjust(pvals_inter$Pval.wo.weights,method='BH')
pvals_inter["Fdr.w.weights"] = p.adjust(pvals_inter$Pval.w.weights,method='BH')


kable(pvals_inter, format = "html") %>%
  kable_styling(bootstrap_options = "striped", full_width = F) 


```
After applying the multiple comparison adjustment, none of the interaction effect terms were significant.
